---
title: "02.analyse.convolvulus.microniche"
author: "Guillaume Papuga"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading data
# I. Micro-niche data ##> to be updated
```{r}
# load raw data
d.flore = read.csv(here::here("data","raw", "flore.convol_complet.csv"), sep = ";", dec = ",") # plant community
d.niche = read.csv(here::here("data","raw", "micro_niche.csv"), sep = ";", dec = ",") # micro niche
d.station = read.csv(here::here("data","raw", "info_station.csv"), sep = ";", dec = ",") # station

# calculate plat community index 

# join d.flore and d.niche
d.microniche
nb.row = nrow(d.microniche)

# write data in the "processed" folder
write.table (x = d.microniche, here::here("data","processed","d.microniche.txt"), sep = ";", dec = ".")
```
The *niche* dataset contains `r nb.row` populations.


```{r cars}
summary(cars)
```

## Microniche analysis

Test for the niche analysis.

```{r pressure, echo=FALSE}
plot(pressure)
```

We'll see if it works. 
Second test now.
An that's a third test. 
And a fourth test. 


And now, this is a first test to see if the push-command works (hopefuly yes)


Woow second test

That's a whole new test to add stuff from another session. (the github buttons does not appear)